---
title: "dashboard"
date: "11/29/2019"
output: 
  flexdashboard::flex_dashboard:
    source_code: embed
    vertical_layout: scroll
    theme: yeti
---

``` {js}
// inverse color of navigation bar
$('.navbar-inverse').removeClass('navbar-inverse').addClass('navbar-default');
```

```{r setup, include=FALSE}
# Load necessary packages. 
library(flexdashboard) # dashboard
library(highcharter) # interactive data visualizations
library(plotly) # interactive data visualizations
library(viridis) # color gradients
library(tidyverse) # metapackge
library(countrycode) # convert country names/codes
library(rjson) # JSON reader
library(crosstalk) # provide interactivity for HTML widgets
library(DT) # display data tables
```

```{r include=FALSE}
# read in data
data = read.csv('./_site/data/who_suicide_statistics.csv') %>%
  filter(country != 'Dominica',
         country != 'Saint Kitts and Nevis')
data = data[complete.cases(data), ]
         
# fix the names of some of the countries in our data to match the country names used by our map later on so that they'll be interpreted and displayed 
data = data %>%
  mutate(country = fct_recode(country, "The Bahamas" = "Bahamas"),
         country = fct_recode(country, "Cape Verde" = "Cabo Verde"),
         country = fct_recode(country, "South Korea" = "Republic of Korea"),
         country = fct_recode(country, "Russia" = "Russian Federation"),
         country = fct_recode(country, "Republic of Serbia" = "Serbia"),
         country = fct_recode(country, "United States of America" = "United States"))

# reorder levels of age to be in chronological order 
data$age = factor(data$age, levels = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"))
```

```{r include=FALSE}
# create a custom theme for the plots 
custom_theme = hc_theme(
  colors = c('#5CACEE', 'green', 'red'),
  chart = list(
         backgroundColor = '#FAFAFA', 
         plotBorderColor = "black"),
  xAxis = list(
         gridLineColor = "E5E5E5", 
         labels = list(style = list(color = "#333333")), 
         lineColor = "#E5E5E5", 
         minorGridLineColor = "#E5E5E5", 
         tickColor = "#E5E5E5", 
         title = list(style = list(color = "#333333"))), 
  yAxis = list(
         gridLineColor = "#E5E5E5", 
         labels = list(style = list(color = "#333333")), 
         lineColor = "#E5E5E5", 
         minorGridLineColor = "#E5E5E5", 
         tickColor = "#E5E5E5", 
         tickWidth = 1, 
         title = list(style = list(color = "#333333"))),   
  title = list(style = list(color = '#333333', fontFamily = "Lato")),
  subtitle = list(style = list(color = '#666666', fontFamily = "Lato")),
  legend = list(
         itemStyle = list(color = "#333333"), 
         itemHoverStyle = list(color = "#FFF"), 
         itemHiddenStyle = list(color = "#606063")), 
  credits = list(style = list(color = "#666")),
  itemHoverStyle = list(color = 'gray'))
```

<style>
.colored {
  background-color: #FAFAFA;
}
</style>


Worldwide
=======================================================================

 Column {.tabset .tabset-fade data-width=700 .colored }
-----------------------------------------------------------------------

### Worldwide suicides {.no-padding}
```{r fig.height=5}
# create tibble for our line plot  
overall_tibble = data %>%
  select(year, suicides_no, population) %>%
  group_by(year) %>%
  summarise(suicide_capita = round((sum(suicides_no)/sum(population))*100000, 2)) 

# create a line plot
highchart() %>% 
    hc_add_series(overall_tibble, hcaes(x = year, y = suicide_capita, color = suicide_capita), type = "line") %>%
    hc_tooltip(crosshairs = TRUE, borderWidth = 1.5, headerFormat = "", pointFormat = paste("Year: <b>{point.x}</b> <br> Suicides: <b>{point.y}</b>")) %>%
    hc_title(text = "Worldwide suicides by year") %>% 
    hc_subtitle(text = "1979-2016") %>%
    hc_xAxis(title = list(text = "Year")) %>%
    hc_yAxis(title = list(text = "Suicides per 100K people"),
             allowDecimals = FALSE,
             plotLines = list(list(
                    color = "black", width = 1, dashStyle = "Dash", 
                    value = mean(overall_tibble$suicide_capita),
                    label = list(text = "Mean = 13.09", 
                                 style = list(color = "black", fontSize = 11))))) %>%
    hc_legend(enabled = FALSE) %>% 
    hc_add_theme(custom_theme)
```

### Worldwide suicides by Gender {.no-padding}
```{r fig.height=5.14}
# create tibble for sex so it can be used to create our line plot 
sex_tibble = data %>%
  select(year, sex, suicides_no, population) %>%
  group_by(year, sex) %>%
  summarise(suicide_capita = round((sum(suicides_no)/sum(population))*100000, 2))

# pick color for gender
sex_color = c("#EE6AA7", "#87CEEB") # baby blue & pink

# create line plot
highchart() %>% 
    hc_add_series(sex_tibble, hcaes(x = year, y = suicide_capita, group = sex), type = "line", color = sex_color) %>%
    hc_tooltip(crosshairs = TRUE, borderWidth = 1.5, headerFormat = "", pointFormat = paste("Year: <b>{point.x}</b> <br>","Gender: <b>{point.sex}</b><br>", "Suicides: <b>{point.y}</b>")) %>%
    hc_title(text = "Worldwide suicides by Gender") %>% 
    hc_subtitle(text = "1979-2016") %>%
    hc_xAxis(title = list(text = "Year")) %>%
    hc_yAxis(title = list(text = "Suicides per 100K people"),
             allowDecimals = FALSE,
             plotLines = list(list(
                    color = "black", width = 1, dashStyle = "Dash",
                    value = mean(overall_tibble$suicide_capita),
                    label = list(text = "Mean = 13.09", 
                                 style = list(color = 'black', fontSize = 11))))) %>% 
    hc_add_theme(custom_theme)
```

### Worldwide suicides by Age {.no-padding}
```{r fig.height=5.13}
# create tibble for age so it can be used to create our line plot 
age_tibble = data %>%
  select(year, age, suicides_no, population) %>%
  group_by(year, age) %>%
  summarise(suicide_capita = round((sum(suicides_no)/sum(population))*100000, 2))

# pick color for graph
age_color = rev(plasma(6))

# create a line plot
highchart() %>% 
    hc_add_series(age_tibble, hcaes(x = year, y = suicide_capita, group = age), type = "line", color = age_color) %>%
    hc_tooltip(crosshairs = TRUE, borderWidth = 1.5, headerFormat = "", pointFormat = paste("Year: <b>{point.x}</b> <br>","Age: <b>{point.age}</b><br>", "Suicides: <b>{point.y}</b>")) %>%
    hc_title(text = "Worldwide suicides by Age") %>% 
    hc_subtitle(text = "1979-2016") %>%
    hc_xAxis(title = list(text = "Year")) %>%
    hc_yAxis(title = list(text = "Suicides per 100K people"),
             allowDecimals = FALSE,
             plotLines = list(list(
                    color = "black", width = 1, dashStyle = "Dash",
                    value = mean(overall_tibble$suicide_capita),
                    label = list(text = "Mean = 13.09", 
                                 style = list(color = 'black', fontSize = 11))))) %>% 
    hc_add_theme(custom_theme)
```


Column {data-width=300}
-----------------------------------------------------------------------

### Suicides per 100K <font size='1'> (1979-2016) </font>
```{r fig.height=0.8}
# grab worldwide number of suicides per 100K people from the data
total_suicides = round(mean(overall_tibble$suicide_capita), 2)

# create value box
valueBox(total_suicides, icon = "fa-plus", color = 'firebrick')
```

### Worldwide suicides by Gender {.no-title .no-padding .colored }
```{r fig.height=1.9}
# make a tibble of suicide by sex
pie_sex = data %>%
  select(sex, suicides_no, population) %>%
  group_by(sex) %>%
  summarise(suicide_capita = round((sum(suicides_no)/sum(population))*100000, 2))
  
# create pie chart for sex 
highchart() %>% 
  hc_add_series(pie_sex, hcaes(x = sex, y = suicide_capita, 
                               color = sex_color), type = "pie") %>%
  hc_tooltip(borderWidth = 1.5, headerFormat = "", pointFormat = paste("Gender: <b>{point.sex} ({point.percentage:.1f}%)</b> <br> Suicides per 100K: <b>{point.y}</b>")) %>%
  hc_title(text = "<b>Worldwide suicides by Gender</b>", style = (list(fontSize = '14px'))) %>% 
  hc_subtitle(text = "1979-2016", style = (list(fontSize = '10px'))) %>%
  hc_plotOptions(pie = list(dataLabels = list(distance = 5, 
                            style = list(fontSize = 10)), 
                            size = 130)) %>% 
  hc_add_theme(custom_theme)
```

### Worldwide suicides by Age {.no-title .no-padding .colored }
```{r fig.height=1.9}
# create a tibble of suicide by age
pie_age = data %>%
  select(age, suicides_no, population) %>%
  group_by(age) %>%
  summarise(suicide_capita = round((sum(suicides_no)/sum(population))*100000, 2)) %>%
  arrange(suicide_capita)

# create pie chart for age
highchart() %>% 
  hc_add_series(pie_age, hcaes(x = age, y = suicide_capita, 
                               color = age_color), type = "pie") %>%
  hc_tooltip(borderWidth = 1.5, headerFormat = "", pointFormat = paste("Age: <b>{point.age} ({point.percentage:.1f}%)</b> <br> Suicides per 100K: <b>{point.y}</b>")) %>%  
  hc_title(text = "<b>Worldwide suicides by Age</b>", style = (list(fontSize = '14px'))) %>% 
  hc_subtitle(text = "1979-2016", style = (list(fontSize = '10px'))) %>%
  hc_plotOptions(pie = list(dataLabels = list(distance = 5, 
                            style = list(fontSize = 10)), 
                            size = 130)) %>% 
  hc_add_theme(custom_theme)
```

Continents 
========================================================================

Column {data-width=375}
-----------------------------------------------------------------------

```{r include=FALSE}
# create new column in data for continent 
# use countrycode() to extract continents from country names
data$continent = countrycode(sourcevar = data$country,
                              origin = "country.name",
                              destination = "continent")

# reclassify countries that have been coded as 'Americas' into 'North America' and 'South America' by countrycode()
south_america = c('Argentina', 'Brazil', 'Chile', 'Colombia', 'Ecuador', 'Guyana', 'Paraguay', 'Suriname', 'Uruguay')

data$continent[data$country %in% south_america] = 'South America'
data$continent[data$continent == 'Americas'] = 'North America'
```

### Suicides by continent and Gender {.no-title .no-padding .colored }
```{r fig.height=2.35}
# create a tibble for continent and sex
continent_sex_tibble = data %>%
  select(continent, sex, suicides_no, population) %>%
  group_by(continent, sex) %>%
  summarize(suicide_capita = round((sum(suicides_no)/sum(population))*100000, 2))

# create histogram of suicides by continent
highchart() %>%
hc_add_series(continent_sex_tibble, hcaes(x = continent, y = suicide_capita, group = sex), type = "column")  %>% 
    hc_colors(colors = sex_color) %>%
    hc_title(text = "Suicides by continent and <b>Gender</b>", style = (list(fontSize = '14px'))) %>% 
    hc_subtitle(text = "1979-2016") %>%
    hc_tooltip(borderWidth = 1.5, pointFormat = paste("Gender: <b> {point.sex} </b> <br> Suicides: <b>{point.y}</b>")) %>%
    hc_xAxis(categories = c("Africa", "Asia", "Europe", "North <br> America", "Oceania", "South <br> America"), labels = list(style = list(fontSize = 8))) %>%
    hc_yAxis(labels = list(style = list(fontSize = 10)),
             title = list(text = "Suicides per 100K people",
             style = list(fontSize = 10)),
        plotLines = list(
          list(color = "black", width = 1, dashStyle = "Dash", 
               value = mean(overall_tibble$suicide_capita),
               label = list(text = "Mean = 13.09", style = list(color = "black", fontSize = 6))))) %>%     
    hc_legend(verticalAlign = 'top', enabled = FALSE) %>% 
    hc_add_theme(custom_theme)
```

### Suicides by continent and Age {.no-title .no-padding .colored }
```{r fig.height=2.35}
# create a tibble for continent and sex
continent_age_tibble = data %>%
  select(continent, age, suicides_no, population) %>%
  group_by(continent, age) %>%
  summarize(suicide_capita = round((sum(suicides_no)/sum(population))*100000, 2)) 

# create histogram of suicides by continent.
highchart() %>%
hc_add_series(continent_age_tibble, hcaes(x = continent, y = suicide_capita, group = age), type = "column")  %>% 
    hc_colors(colors = age_color) %>%
    hc_title(text = "Suicides by continent and <b>Age</b>", style = (list(fontSize = '14px'))) %>% 
    hc_subtitle(text = "1979-2016") %>%
    hc_tooltip(borderWidth = 1.5, pointFormat = paste("Age: <b> {point.age} </b> <br> Suicides: <b>{point.y}</b>")) %>%
    hc_xAxis(categories = c("Africa", "Asia", "Europe", "North <br> America", "Oceania", "South <br> America"), labels = list(style = list(fontSize = 8))) %>%
    hc_yAxis(labels = list(style = list(fontSize = 10)),
             title = list(text = "Suicides per 100K people",
                          style = list(fontSize = 10)),
        plotLines = list(
          list(color = "black", width = 1, dashStyle = "Dash", 
               value = mean(overall_tibble$suicide_capita),
               label = list(text = "Mean = 13.09", style = list(color = "black", fontSize = 6))))) %>%    
    hc_legend(verticalAlign = 'top', enabled = FALSE) %>% 
    hc_add_theme(custom_theme)
```


Column {data-width=625}
-----------------------------------------------------------------------

### Suicides by continent {.no-title .no-padding .colored }
```{r}
# import continent map data from the internet. 
map_data = download_map_data("custom/world-continents")

# create a tibble for continent
continent_tibble = data %>%
  select(continent, suicides_no, population) %>%
  group_by(continent) %>%
  summarize(suicide_capita = round((sum(suicides_no)/sum(population))*100000, 2)) %>%
  arrange(suicide_capita)

# create continent map with suicide data 
highchart() %>%
hc_add_series_map(map_data, continent_tibble, value = "suicide_capita", joinBy = c('name','continent'), name = "Suicides (per 100K people)")  %>% 
    hc_add_series(continent_tibble, hcaes(x = continent, y = suicide_capita, color = suicide_capita), type = "pie", name = 'Suicides (per 100K people)')  %>% 
    hc_colorAxis(stops = color_stops()) %>% 
    hc_title(text = "Suicides by Continent") %>% 
    hc_subtitle(text = "1979-2016") %>%
    hc_tooltip(borderWidth = 1.5, valueSuffix = '') %>%
    hc_plotOptions(
    pie = list(center = c('10%', '80%'), size = 110, dataLabels = list(enabled = FALSE))) %>% 
    hc_add_theme(custom_theme)
```

Countries
========================================================================

Column {.tabset .tabset-fade data-width=400 .colored }
-----------------------------------------------------------------------

### By country {.no-padding}
```{r fig.height=4.55}
# create tibble for overall suicides by country
country_bar = data %>%
  select(country, suicides_no, population) %>%
  group_by(country) %>%
  summarise(suicide_capita = round((sum(suicides_no)/sum(population))*100000, 2)) %>%
  arrange(desc(suicide_capita))

# create interactive bar plot
highchart() %>%
    hc_add_series(country_bar, hcaes(x = country, y = suicide_capita, color = suicide_capita), type = "bar")  %>% 
    hc_tooltip(borderWidth = 1.5, 
               pointFormat = paste("Suicides: <b>{point.y}</b>")) %>%
    hc_legend(enabled = FALSE) %>%
    hc_title(text = "Suicides by country") %>% 
    hc_subtitle(text = "1979-2016") %>%
    hc_xAxis(categories = country_bar$country, 
             labels = list(step = 1),
             min = 0, max = 25,
             scrollbar = list(enabled = TRUE)) %>%
    hc_yAxis(title = list(text = "Suicides per 100K people")) %>%
    hc_plotOptions(bar = list(stacking = "normal", 
                              pointPadding = 0, groupPadding = 0, borderWidth = 0.5)) %>% 
    hc_add_theme(custom_theme)
```

### By gender {.no-padding}
```{r fig.height=4.65}
# create tibble for suicide by countries and sex 
country_bar_sex = data %>%
  select(country, sex, suicides_no, population) %>%
  group_by(country, sex) %>%
  summarise(suicide_capita = round((sum(suicides_no)/sum(population))*100000, 2))

country_tibble = data %>%
  select(country, suicides_no, population) %>%
  group_by(country) %>%
  summarise(suicide_capita = round((sum(suicides_no)/sum(population))*100000, 2)) 

# create bar chart of suicide by sex
highchart() %>%
    hc_add_series(country_bar_sex, hcaes(x = country, y = suicide_capita, group = sex), type = "bar", color = sex_color)  %>% 
    hc_tooltip(borderWidth = 1.5, pointFormat = paste("Gender: <b>{point.sex} ({point.percentage:.1f}%)</b> <br> Suicides per 100K: <b>{point.y}</b>")) %>%
    hc_legend(enabled = TRUE, colorByPoint = TRUE) %>%
    hc_title(text = "Suicides by country and gender") %>% 
    hc_subtitle(text = "1979-2016") %>%
    hc_xAxis(categories = country_tibble$country,
             labels = list(step = 1),
             min = 0, max = 25,
             scrollbar = list(enabled = TRUE)) %>%
    hc_yAxis(title = list(text = "Percentage of total suicides")) %>%
    hc_plotOptions(bar = list(stacking = "percent", 
                              pointPadding = 0, groupPadding = 0, borderWidth = 0.4)) %>% 
    hc_add_theme(custom_theme)
```

### By age {.no-padding}
```{r fig.height=4.65}
# create tibble for suicide by countries and age 
country_bar_age = data %>%
  select(country, age, suicides_no, population) %>%
  group_by(country, age) %>%
  summarise(suicide_capita = round((sum(suicides_no)/sum(population))*100000, 2))

# create interactive bar plot
highchart() %>%
    hc_add_series(country_bar_age, hcaes(x = country, y = suicide_capita, group = age), type = "bar", color = age_color)  %>% 
    hc_tooltip(borderWidth = 1.5, pointFormat = paste("Age: <b>{point.age} ({point.percentage:.1f}%)</b> <br> Suicides per 100K: <b>{point.y}</b>")) %>%
    hc_title(text = "Suicides by country and age") %>% 
    hc_subtitle(text = "1979-2016") %>%
    hc_xAxis(categories = country_tibble$country,
             labels = list(step = 1),
             min = 0, max = 25,
             scrollbar = list(enabled = TRUE)) %>%
    hc_yAxis(title = list(text = "Percent of total suicides")) %>%
    hc_plotOptions(bar = list(stacking = "percent", 
                              pointPadding = 0, groupPadding = 0, borderWidth = 0.5)) %>% 
    hc_add_theme(custom_theme)
```


Column {data-width=600}
-----------------------------------------------------------------------

### Suicides by country {.no-title .no-padding .colored }
```{r fig.height=4.9}
# create a tibble with suicide per capita by country for 1985-2015
country_tibble = data %>%
  select(country, suicides_no, population) %>%
  group_by(country) %>%
  summarize(suicide_capita = round((sum(suicides_no)/sum(population))*100000, 2))

# create interactive world map
highchart() %>%
hc_add_series_map(worldgeojson, country_tibble, value = "suicide_capita", joinBy = c('name','country'))  %>% 
    hc_colorAxis(stops = color_stops()) %>% 
    hc_title(text = "Suicides by Country") %>% 
    hc_subtitle(text = "1979-2016") %>%
    hc_tooltip(borderWidth = 1.5, headerFormat = "", valueSuffix = " suicides (per 100K people)") %>% 
    hc_add_theme(custom_theme)
```

Search {data-icon="fa-search"}
=======================================================================

Column {data-width=250}
-----------------------------------------------------------------------

### Filters {.no-title .colored }

**Pick filters here:**
```{r}
# create tibble for our line plot 
country_year_tibble = data %>%
  select(country, year, suicides_no, population) %>%
  group_by(country, year) %>%
  summarise(suicide_capita = round((sum(suicides_no)/sum(population))*100000, 2)) 

# create shared data that will be used to link filters, data table, and line plot 
shared_data = SharedData$new(country_year_tibble, group = 'hello')

# create filter for year and country
# adjust the DT datatable and PLOTLY plot

filter_slider("year", "Year", shared_data, ~ year, step = 1)
filter_select("country", "Country", shared_data, ~ country, allLevels = TRUE, multiple = TRUE)
```

***

**Data table**
```{r}
# create datatable
datatable(shared_data,
          rownames = FALSE,
          colnames = c('Country', 'Year', 'Suicides /100K'),
          class = 'cell-border stripe',
          extensions = "Scroller",
          options=list(deferRender = FALSE, 
                       scrollY = 200, 
                       scrollCollapse = TRUE,
                       scroller = TRUE,
                       dom = 't'))
```


Column {data-width=750}
-----------------------------------------------------------------------

### Suicides by country {.no-title .colored }
```{r}
# set a random seed to reproduce the random sample of colors we will use for our next graph 
set.seed(20191029)

# create line graph 
plot_ly(shared_data, x = ~ year, y = ~ suicide_capita, 
       color = ~country, colors = sample(colours(), 120),
       type = 'scatter', mode = 'lines',
       hoverinfo = 'text', text = ~ paste("Country: ", country, '<br>Year: ', year, "<br>Suicides: ", suicide_capita)) %>%
       layout(showlegend = FALSE,
              title = "Suicide by country",
              xaxis = list(title = "Year"),
              yaxis = list(title = "Suicides per 100K people")) %>%
       layout(plot_bgcolor = 'transparent') %>% 
       layout(paper_bgcolor = 'transparent') %>% 
       add_markers() %>% 
       highlight("plotly_click")
```